Type Error Telemetry at Scale: Pseudonymized and Stochiastic

### 2023-09-14 programming23/paper.tex
intro pass

\title{Millions of Type Errors}
\subtitle{Impersonal Telemetry to Measure User Experience}
% lean, lite, modest, non-intrusive, private, ...

@@

  \paragraph{Importance}
  %% Why does this work matter?
  Beyond the immediate benefits to Luau,
  our findings about types and type errors have implications
  for adoption and ergonomics in other gradually typed languages.
  %% mention Elixir, Verse?
  FILL specifics.
  Our telemetry design is of broad interest, as it gathers useful
  data without exposing personal information.

@@

{Roblox experiences} are scripted using the 
{Luau} programming language~\cite{luau-lang.org},
an extension of {Lua~5.1~\cite{lua}}.
The main extension is the addition of a static type system, which uses
type inference to synthesize types for user code. These types
are used primarily in type-driven tooling such as autocomplete
and API documentation~\cite{luau-autocomplete},
but creators can also opt in to receiving type error reports.

As discussed in~\cite{bfj-hatra-2021},
the goals of the {Luau} type system are rather different from
a traditional type system, which focuses on compilation and memory safety.
{Luau} has a diverse user community, ranging from
students in code camps to professional development studios. These
creators have quite different needs, with different emphases on
enabling rapid creation and ensuring software quality.
{Luau} therefore supports three typing modes---strict types,
relaxed (nonstrict) types, and no types---and lets users switch
between modes gradually~\cite{st-sfp-2006,tfffgksst-snapl-2017}, one module at a time.

In this paper, we investigate methods for measuring the effectiveness
of the {Luau} type system in the development of {Roblox} scripts.
In comparison to prior work~(\cref{s:related}), which is either small in scale
or depends on personally identifiable information~(PII),
we performed a large-scale study using pseudonymized \emph{telemetry}.

@@

The telemetry data we analyzed does not contain any PII:
no source code;
no source code locations;
no error messages (which may contain source code);
no record of the creator's identity, locale, or IP address;
and no information about what creation the data came from.
Telemetry records are correlated by session, using a pseudonymized
session identifier.

@@

At a higher level, this paper is the first to use telemetry 
to study a typechecker.
It thus represents a step toward data-driven language design,
informed by many users' actual practice.
Our data captures \textbf{XX} sessions
for a total of \textbf{YY} hours
and \textbf{ZZ} type errors.
By contrast to typical qualitative methods (e.g., surveys and interviews), it
is not restricted to users' \emph{perceptions} about their work and it is not
limited to a small number of users.

@@

other Nasko papers:
https://dblp.org/pid/83/1412.html

>>>>

\title{Type Error Telemetry at Scale: Pseudonymized and Stochiastic}
% lean, lite, modest, non-intrusive, private, ...
% {Millions of Type Errors} = 72M forced strict, 1/2M type error is disappointing!
%% \subtitle{Impersonal Telemetry to Measure User Experience}

@@

  \paragraph{Importance}
  %% Why does this work matter?
  Luau is the first large-scale combination of gradual types and semantic
  subtyping.
  Beyond the immediate benefits to Luau,
  our findings about types and type errors have implications
  for adoption and ergonomics in other gradual languages.
  %% mention Elixir, Verse?
  FILL specifics.
  Our telemetry design is of broad interest, as it gathers useful
  data without exposing personal information.

@@

{Roblox experiences} are scripted using the 
{Luau} programming language~\cite{luau-lang.org},
an extension of {Lua~5.1~\cite{lua}}.
The main extension is the addition of a static type system, which uses
type inference to synthesize types for user code on the fly.
These types are used primarily in type-driven tooling such as autocomplete and
API documentation~\cite{luau-autocomplete}, but creators can also opt in to
receiving type error reports.

As discussed in~\cite{bfj-hatra-2021},
the goals of the {Luau} type system are rather different from
a traditional type system, which focuses on compilation and memory safety.
{Luau} has a diverse user community, ranging from
students in code camps to professional development studios. These
creators have quite different needs, with different emphases on
enabling rapid creation and ensuring software quality.
{Luau} therefore supports three typing modes---\mstrict{} types,
relaxed (\mnonstrict{}) types, and no types (\mnocheck{})---and lets users switch
between modes gradually~\cite{st-sfp-2006,tfffgksst-snapl-2017}, one module at a time.

In this paper, we investigate methods for measuring the effectiveness
of the {Luau} type system in the development of {Roblox} scripts.
We want to collect measurements at a large scale, with thousands
of participants, and we must do so in a way that protects creators' privacy.
In comparison to prior work~(\cref{s:related}), which
with few exceptions~\cite{zhlbr-cc-2020,zhlbr-oopsla-2020,hlzbr-ecoop-2021} is either small in scale
or collects personally identifiable information~(PII),
we performed a large-scale study using pseudonymized \emph{telemetry}.

@@

The telemetry data we analyzed does not contain any PII.
It has no source code;
 no source locations;
no error messages (which may contain source code);
no error messages;
no record of the creator's identity, locale, or IP address;
and no information about what creation the data came from.
Telemetry records are correlated by session, using a pseudonymized
session identifier.

@@

At a higher level, this paper is the first to use telemetry 
to study a typechecker.
It thus represents a step toward data-driven language design,
informed by many users' actual practice.
Our data captures over 340,000 sessions
that occured between February and April 2023
and includes 72 million type analysis errors.
By contrast to typical qualitative methods (e.g., surveys and interviews), it
is not restricted to users' \emph{perceptions} about their work and it is not
limited to a small number of users.

@@

Nasko~\cite{zhlbr-cc-2020,zhlbr-oopsla-2020,hlzbr-ecoop-2021}
